---
title: "CodesMaster"
author: "Jonathan Slaney"
date: "19/02/2020"
output: html_document
---

### Write a file

```{r}
write.csv(filename, file = "/Users/jslaney/filename.csv")
```

### Read in and combine (bind) dataframes

```{r}
setwd("C:/Users/jslan/Desktop/veritas.w2")
w2new <- read.csv("health_2vicnew_main_bfd440e.csv")
w2ret <- read.csv("health_2vic_main_bfd440e.csv")
w1 <- read.csv("health_1vic_main_bfd440e.csv")
#add wav

d <- bind_rows(w2new, w2ret, w1)

## Another example
merged_data <- bind_rows(p1, p2)
view(merged_data)
summary(merged_data)

## Read in and rename Cols

p1_hip <- read.csv("/users/jslaney/desktop/graduate_studies/hkr6130/accelerometer/p1_right hip.csv", header = FALSE, skip = 100, col.names = c("time", "x_axis", "y_axis", "z_axis", "lux", "button", "temp"))

head(p1_hip)
```

### Use purrr:map to read in all csv's in a folder

```{r}
library(readr)
awfiles_df <- NULL
awfiles <- list.files(path = "C:/Users/jslan/Desktop/aw_data_folder", pattern = "*.csv", full.names= T)

awfiles %>%
    purrr::map(function(i){
      ndf <- read.csv(i, header = TRUE) 
      awfiles_df <<- bind_rows(awfiles_df, ndf)
    })

summary(awfiles_df)

## Compute mean and sum data for each participant (group_by(Username))
summary_table <- awfiles_df %>%
                  group_by(Username) %>% 
                    summarize(
                      heart_mean = mean(Heart, na.rm = TRUE),
                      calories_summ = sum(Calories, na.rm = TRUE),
                      steps_summ = sum(Steps, na.rm = TRUE), 
                      distance_sum = sum(Distance, na.rm = TRUE)
                    )
summary_table
```

### Basic 2 variable plot

```{r}
scatter <- ggplot(cchs, aes(x = hwtghtm, y = hwtgwtk)) + 
            geom_point() + 
              xlab("Height in Meters") + 
              ylab("Weight in Kilograms")
plot(scatter)
```

### Facet wrap layer for ggplot of continous variables

```{r}
ab <- ggplot(d, aes(x = d$transp_bikes_adults)) + 
  geom_histogram(binwidth = 1, fill="#76D24A", na.rm = TRUE) + 
  xlab("Number of bicycles for adults") + 
  scale_x_continuous(breaks = c(5, 10, 15, 20)) 

ab + facet_wrap(vars(wave_id))
  
summary(d$transp_bikes_adults)
```

### Mutate a column into specific categories

```{r}
# Calculate BMI and store in new var
cchs <- cchs %>%
  mutate(weight = ifelse(hwtgwtk == 999.96 | hwtgwtk == 999.99, NA, hwtgwtk),
         bmi = weight/height^2)

# Categorize new var contents
cchs <- cchs %>%
          mutate(bmi_category = case_when(
            bmi < 18.5 ~ "underweight",
            bmi >=30 & bmi <999 ~ "obese",
            bmi >=25 & bmi <30 ~ "overweight",
            bmi >=18.5 & bmi <25 ~ "normal weight",
            TRUE ~ "other")
            )

## Summarize the entire df
summary(cchs)

## Another example

steps<-aw_data$steps

aw_data <- aw_data %>%
	mutate(steps_type = case_when(
		steps == 0 ~ "No steps",
		steps >=100 & steps <200 ~ "vigorous",
		steps >=0 & steps <100 ~ "moderate",
		TRUE ~ "other"
	))
  
view(aw_data)
```

### Recode col contents names

```{r}
cchs <- cchs %>%
	mutate(province = case_when(
		geogprv == 10 ~ "NFLD & LAB",
		geogprv == 11 ~ "PEI",
		geogprv == 12 ~ "NOVA SCOTIA",
		geogprv == 13 ~ "NEW BRUNSWICK",
		geogprv == 24 ~ "QUEBEC",
		geogprv == 35 ~ "ONTARIO",
		geogprv == 46 ~ "MANITOBA",
		geogprv == 47 ~ "SASKATCHEWAN",
		geogprv == 48 ~ "ALBERTA",
		geogprv == 59 ~ "BRITISH COLUMBIA",
		geogprv == 60 ~ "YUKON/NWT/NUNA",
		geogprv == 96 ~ "NOT APPLICABLE",
		geogprv == 97 ~ "DON'T KNOW",
		geogprv == 98 ~ "REFUSAL",
		TRUE ~ "NOT STATED"
	))
```

### Another ay to rename observations of a var

```{r}
d$bike_safety_f <- revalue(as.factor(d$bike_safety), c("1" = "Very safe", "2" = "Somewhat safe", "3" = "Neither safe nor unsafe", "4" = "Somewhat dangerous", "5" = "Very dangerous"))
```

### Create tables of one variable totals, grouped by another variable and diplayed side by side (facet_wrap)

```{r}

t_1 <- d %>%
          group_by(wave_id, bike_safety_f) %>%
            summarise(n = n()) %>%
            mutate(pct = n / sum(n))

kable(t_1)  %>%   kable_styling(bootstrap_options = "striped", full_width = T, position = "left")

```

### Create plot of one variable, grouped by another variable (ID) + facet_wrapped

```{r}
p <- ggplot(d, aes(bike_safety,  fill = bike_safety_f)) + theme(axis.text.x = element_text(angle=90, vjust = .6)) + 
  geom_bar() +
  scale_fill_manual(values = INTERACTPalette3) +
  guides(fill=FALSE) +
      ylab("Count") +
      xlab("Perception of bicycle risk") +
      ggtitle("Overall, how safe do you think cycling is in your city?") + 
  facet_wrap(~ wave_id)
plot(p)
```

### Remove unwanted values with DPLYR::mutate() 

```{r}
cchs <- cchs %>%
  mutate(height = ifelse(hwtghtm == 9.996 | hwtghtm == 9.999, NA, hwtghtm))
```

### GROUP_BY() SUMMARIZE() <- get the mean and SD of BMI for each category in one code chunk. 

```{r}
bmi_prov <- cchs %>%
  group_by(province) %>%
    summarise(mean_BMI = mean(bmi, na.rm=TRUE),
              sd_BMI = sd(bmi, na.rm=TRUE),
              total = n())

## Another example

summary_table <- aw_data %>%
                  group_by(steps_type) %>% 
                  summarize(
                  heart_mean = mean(heart_clean, na.rm = TRUE),
                  calories_mean = mean(calories, na.rm = TRUE)
                    )
view(summary_table)
```

### How to Subset data

```{r}
newfoundland <- cchs %>%
  filter(province == "NFLD & LAB")  

sask <- cchs %>%
  filter(province == "SASKATCHEWAN")

# stack the data bind_rows()

sask_nl <- bind_rows(sask, newfoundland)
table(sask_nl$province)
```

### How to merge data

```{r}
### Select the first set of data
first <- dataframe %>%
  select(wave_id, bike_children)

### Select the second set of data 
second <- dataframe %>%
  select(wave_id, bike_safety)

### Join together

join_full_data <- full_join(first, second, by="wave_id")
head(join_full_data)
```

### Accel data - Aggregate accelerometer data by second and calculate the gravity subtracted vector magnitude, and the standard deviation for each axis (x, y, z,)

```{r}
accel_data$time1 <- str_sub(accel_data$time, 1, str_length(accel_data$time)-4) 
  ### Removes all milliseconds

accel_data$time2 <- ymd_hms(accel_data$time1)
  ### Create a POSIXct formated variable

accel_data$hour <- hour(accel_data$time2)
  ### Create a hour variable

accel_data$minute <- minute(accel_data$time2)
  ### Create a minute variable

accel_data$second <- second(accel_data$time2)
  ### Create a second variable

## Create Vector Magnitude variable
vec_mag <- sqrt(accel_data$x_axis^2+accel_data$y_axis^2+accel_data$z_axis^2)

## Gravity subtracted vector magnitude
vec_mag_g <- abs(sqrt(accel_data$x_axis^2+accel_data$y_axis^2+accel_data$z_axis^2)-1)

# AGGREGATE

accel_sec <- accel_data %>%
  group_by(id, wear_loc, hour, minute, second) %>%
  summarise(
         time = first(time),
         m_x_axis = mean(x_axis),
         m_y_axis = mean(y_axis),
         m_z_axis = mean(z_axis),
         vec_mag = mean(vec_mag),
         vec_mag_g = first(vec_mag_g),
         sd_x_axis = sd(x_axis),
         sd_y_axis = sd(y_axis),
         sd_z_axis = sd(z_axis)
        )

```

## Accel data - Using Cut Points Create A New Variable Called Activity

```{r}
accel_sec <- accel_sec %>%
  mutate(activity = case_when(
    vec_mag_g <= 0.19000 ~ "Sedentary",
    vec_mag_g >0.19001 & vec_mag_g <0.31400 ~ "Light",
    vec_mag_g >=0.314001 &  vec_mag_g  <0.9989 ~ "Moderate",
    vec_mag_g >0.999 ~ "Vigorous",
    TRUE ~ "other"
  ))

# Create figure of physical activity

fig1 <- ggplot(accel_sec, aes(x = time, y = vec_mag_g)) + 
          geom_point(alpha = 1/10) + 
            xlab("Time") +
            ylab("Vector Magnitude")

plot(fig1)
```

### Recode the datetime stamp with lubridate()

```{r}
library(lubridate)
aw_data$date_time <- as_date(aw_data$date_time)
aw_data$ymd <- ymd(aw_data$date_time)
table(aw_data$ymd)
```

### Change 0's to NA's in a column

```{r}
heart_clean <- aw_data$heart %>%
  na_if(0)
```

### Remove NA's and calculate mean HR

```{r}
heart_clean %>%
  na_if(0)
  mean(heart_clean,na.rm = TRUE)
  summary(heart_clean)
```

### Select three variables, combine into a dataframe, plot them

```{r}
x<-aw_data$date_time
y1<-aw_data$heart
y2<-aw_data$steps
df<-data.frame(x, y1, y2)
ggplot(df, aes(x, y= value, color = variable)) +
         geom_point(aes(y=y1, col = "y1")) +
         geom_point(aes(y=y2, col = "y2"))
```

### Reading and plotting the generated data per each participant
```{r warning=FALSE}
for (i in 1:length(participants)) {
  current_participant <- participants[i]
  current_path <- paste(current_participant, "/", sep = "")

  participants_data_files <- list.files(path = paste0(path, current_path), pattern = "*.csv$", full.names = F)
  if (length(participants_data_files) == 0)
    next()
  # remove the file include all the others
  participants_data_files <- participants_data_files[-which(grepl(participants_data_files, pattern =  "^.*fitbit.*$"))] 
  
  for (file_index in 1:length(participants_data_files)) {
    
    # read each .csv file
    data <- fread(paste0(path, current_path, participants_data_files[file_index]), data.table = F)
    # convert the DateTime type
    data <- data %>% 
      mutate(
        DateTime = as_datetime(DateTime)
      )
    
    data_plot <- ggplot(data, aes(x = DateTime)) +
    geom_histogram() +
    ggtitle(participants_data_files[file_index])
    
    plot(data_plot)
    ggsave(paste0(path, current_path, participants_data_files[file_index], "_plot.pdf"))
  }
  
}
```

### HEXOSKIN Loop over data, add new columns and retrieve info with str_sub()

```{r}
hexodata_temp <- NULL

hexodata_temp <- dir(path = "/Users/dfuller/Desktop/DatasetsHexoskin/" , full.names = TRUE, pattern = "*.csv", recursive = TRUE) %>%
   purrr::map(function(i){
     dfhx <- read.csv(i, header = TRUE)
     dfhx$city <- stringr::str_sub(i, 42, 45)
     dfhx$participants <- stringr::str_sub(i, 59, 60)
     hexodata_temp <- bind_rows(hexodata_temp, dfhx)
   })

hexo_data <- bind_rows(hexodata_temp)
```

### HEXOSKIN Change col names, create hexo_data and write to csv

```{r}
hexo_data <- hexo_data %>%
  dplyr::rename(
     time = time..s.256., 
     respiratory_rate = breathing_rate..rpm...api.datatype.33.., 
     heart_rate = heart_rate..bpm...api.datatype.19.., 
     minute_ventilation_ml = minute_ventilation..mL.min...api.datatype.36.., 
     cadence = cadence..spm...api.datatype.53.., 
     activity_g = activity..g...api.datatype.49..,
  )
```

### HEXOSKIN Delete useless columns

```{r}
hexo_data <- subset(hexo_data, select = -minute_ventilation..L.min...api.datatype.36..)
hexo_data <- subset(hexo_data, select = -sleep_position..NA...api.datatype.270..)
hexo_data <- subset(hexo_data, select = -X)
hexo_data <- subset(hexo_data, select = -time)
```

### HEXOSKIN Add participant individual level data from separate participant file

```{r}
participants <- read_csv("Hexoskin_participants.csv")

hexo_data_complete <- inner_join(hexo_data, participants, by = "participants")

write.csv(hexo_data_complete, "hexo_data.csv") ## Saving do that I do not break my computer when working on this data! 
```

# STAT codes
